Quantitative Benchmarking for Automatically Generated Conjectures

### 2018-04-24 TEBenchmarkPaper/article.tex
Strengthen motivation

% Lay out the field of TE: motivation, tools, evaluation, problems...
% FIXME: Strengthen motivation:
%  - Automated Scientific Discovery (e.g. Dundee drug discovery "centaur" story)
%  - "Mixed initiative" (social machine?); maybe Kasparov

\emph{Conjecture generation} is the open-ended task of producing conjectures
about a given logical theory which are somehow ``interesting'', and is studied
as a sub-field of \emph{mathematical theory exploration} (MTE).

This has applications wherever we find uses for formal methods: in proof
assistants and their libraries, in mathematics education and research, and in
the specification and verification of software. Although formal methods, and
automated reasoning more generally, have been advocated in mathematics (notably
by Voevodsky~\cite{voevodsky2010univalent} and
Gowers~\cite{ganesalingam2013fully}) and software engineering (for example via
dependently-typed
programming~\cite{McKinna:2006, chlipala2011certified, Xi2003}), they have a
reputation for being difficult and expensive, which has historically limited
their use to high-assurance domains such as aerospace and microprocessor design.
One way to reduce these hurdles is to increase automation, which in the case of
software verification can benefit from the introduction of conjecture generation
tools~\cite{Claessen.Johansson.Rosen.ea:2013}. Another approach is to tailor the
division of labour between humans and machines such that each is playing to
their strengths, which has been referred to as ``centaur
teams''~\cite{harari2017reboot, davenport2015beyond}. In this respect a
conjecture generating tool can search a much larger space of relationships than
a human, whilst leaving the ultimate judgement of which to investigate in more
detail to the user.

Similar tasks are also studied in the domain of Computational \emph{Scientific}
Discovery~\cite{king2004functional,Williams20141289,schmidt2009distilling}, for
example searching for new drugs. The major difference between scientific and
mathematical discovery is that inductive reasoning and experimental testing are,
in principle, more important for the former, although we note that they are also
core principles in the mathematical tools we have investigated (in particular
for ensuring the \emph{plausibility} of conjectures).

We limit ourselves to applications in mathematics, but even here it is difficult
to measure and compare the success of approaches and tools. This is partially
due to their diversity, but also because of the inherent ambiguity of the task
(what counts as ``interesting''?), the different goals emphasised by their
designers and the variety of evaluation methods employed. We attempt to solve
this discrepancy, at least for the foreseeable future, by defining a standard,
unambiguous benchmarking approach with which to compare the conjecture
generation of MTE tools. Our contributions include:

>>>>

% Lay out the field of TE: motivation, tools, evaluation, problems...

\emph{Conjecture generation} is the open-ended task of producing conjectures
concerning a given logical theory which are somehow ``interesting'', and is
studied as a sub-field of \emph{mathematical theory exploration} (MTE). This has
applications wherever we find uses for formal methods: in proof assistants and
their libraries, in mathematics education and research, and in the specification
and verification of software. Although formal methods, and automated reasoning
more generally, have been advocated in mathematics (notably by
Voevodsky~\cite{voevodsky2010univalent} and Gowers~\cite{ganesalingam2013fully})
and software engineering (for example via dependently-typed
programming~\cite{McKinna:2006}), they have a reputation for being difficult and
expensive, which has historically limited their use to high-assurance domains
such as aerospace and microprocessor design.

One path to reducing these barriers is to increase the level of automation.
Tools for automated theorem proving (ATP) have been shown to benefit from the
addition of a conjecture generator, improving their ability to prove properties
relevant to software verification~\cite{Claessen.Johansson.Rosen.ea:2013}.
Another path is to tailor the division of labour between humans and machines
such that each is playing to their strengths, which has been referred to as
``centaur teams''~\cite{harari2017reboot,davenport2015beyond}. Machines can
systematically search a much larger space of possible relationships than humans,
potentially bringing to light novel and surprising connections, whilst allowing
the user to make the ultimate judgement over which of the most promising results
are worth investigating further.

Similar tasks are also studied in the domain of Computational \emph{Scientific}
Discovery~\cite{king2004functional,Williams20141289,schmidt2009distilling}, for
example in the search for new drugs. The major difference between scientific and
mathematical discovery is that inductive reasoning and experimental testing are,
in principle, more important for the former, although we note that they are also
core principles in the mathematical tools we have investigated (in particular
for ensuring the \emph{plausibility} of conjectures).

We limit ourselves to mathematical applications, but even here it is difficult
to measure and compare the success of approaches and tools. This is partially
due to their diversity, but also because of the inherent ambiguity of the task
(what counts as ``interesting''?), the different goals emphasised by their
designers and the variety of evaluation methods employed. We attempt to solve
this discrepancy, at least for the foreseeable future, by defining a standard,
unambiguous benchmarking approach with which to compare the conjecture
generation of MTE tools. Our contributions include:
